---
title: "CDC Child Lead Levels"
author: "Elayna Saint Amour"
date: "December 13, 2021"
output: ioslides_presentation
widescreen: true
runtime: shiny
---
```{r setup, include=FALSE,message=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```

```{r Loading data and fixing it for apps, include=FALSE}
library(readxl)
library(dplyr)
library(shiny)
library(stringr)
library(tidyverse)
library(DT)
MI_2012 <- read_excel("~/STA_518/Preparations/Presentation/data/MI_CountyLevelSummary_2012.xlsx")
MI_2013 <- read_excel("~/STA_518/Preparations/Presentation/data/MI_CountyLevelSummary_2013.xlsx")
MI_2014 <- read_excel("~/STA_518/Preparations/Presentation/data/MI_CountyLevelSummary_2014.xlsx")
MI_2015 <- read_excel("~/STA_518/Preparations/Presentation/data/MI_CountyLevelSummary_2015.xlsx")
MI_2016 <- read_excel("~/STA_518/Preparations/Presentation/data/MI_CountyLevelSummary_2016.xlsx")
#Combined data into one master data set called Total
Total <- bind_rows(MI_2012, MI_2013, MI_2014, MI_2015, MI_2016)
Total <- Total %>% 
    mutate(proportion =  Confirmed/Tested) %>% 
    mutate(percent =  100*proportion) %>% 
    mutate(percent = round(percent,3)) %>% 
#turned subregion into all lower case and removed the word county to match michigan_county data set
    mutate(subregion=tolower(str_replace(subregion, " County","")))%>% 
#I made new variables that capitalized the subregion so the user will see it each county as a capital (ex: Kent)
    mutate(capsubregion1= str_to_title(subregion)) %>% 
    mutate(capsubregion2= str_to_title(subregion))
#Imported map data of the United states
usa <- map_data('usa')
state <- map_data("state")
#subsetting information to from US data to include only michigan and its counties
michigan <- subset(state, region=="michigan")
counties <- map_data("county")
michigan_county <- subset(counties, region=="michigan")
```

## About the Data
-This data was collected by the CDC from 2012-2016.

-Data is for children under the age 16 from Michigan divided by county.

-The variable of interest is the proportion of children with lead levels over 5µg/dL.

## Interactive Map

```{r maps of counties}
knitr::include_app("https://rstudio.gvsu.edu/s/c8e00cfc78a314348f306/p/d0d2c012/")
   
```


## Comparing Counties Over Time

```{r linegraph}
#https://rstudio.github.io/DT/
fluidPage(


        mainPanel(

            textInput(inputId = "subregion1",
                        label = "Input county of interest:",
                        value= "Kent"),

            textInput(inputId = "subregion2",
                      label = "Input county of interest:",
                      value= "Ottawa"),

            plotOutput(outputId = "linegraph")
        )
    )

# Define server ----------------------------------------------------------------

    renderPlot({
      output$linegraph <- renderPlot({
                a <- subset(Total,capsubregion1==input$subregion1)
                b <- subset(Total,capsubregion2==input$subregion2)
                ggplot() +
                    geom_line(a,mapping=aes(x=Year, y=Proportion), color="red") +
                    geom_line(b,mapping=aes(x=Year, y=Proportion), color="blue")


    })
    })

```

## Flint Water Crisis
-Starting in 2014 government officials made the decision to switch water sources from the the Detroit River to the Flint River. 


-Starting in September 2015 doctors reveal they’ve found high levels of lead in Flint children’s bloodstreams and advised Flint to stop using the Flint River as a water system.

-According to Consumer Notice "Nearly 9,000 children were supplied lead-contaminated water for 18 months."

-October 2016 Flint begins using the Detroit River again





## Results Between Counties
-Genesee could not be compared to most of its neighboring counties.

-i.e. Lapeer Oakland and Tuscola.

-Genesee was significantly **higher** Saginaw in 2016 at a 95% confidence level.

```{r 2 prop z tests, include=FALSE}
#This function takes two counties from the same year and tests for differences in proportions
countyproptest <- function(county1, county2, year){
  pairwise_cases <- Total %>%
    select(subregion, Confirmed, Tested,Year) %>%
    filter(Year == {{year}},
           subregion %in% c({{county1}},{{county2}}))

  prop.test(x = pairwise_cases$Confirmed, 
            n = pairwise_cases$Tested,
            alternative = c("two.sided"),
            conf.level = 0.95, correct = TRUE)}
#lapeer had zero cases
#oakland, tuscola did not pass conditions
countyproptest("genesee", "saginaw", 2016)

#This function takes a county and tests for differences in proportions between years
yearproptest <- function(county, year1,year2){
  pairwise_years <-  Total %>% 
    select(subregion,Confirmed, Tested,Year) %>% 
    filter(subregion %in% {{county}},
           Year %in% c({{year1}}, {{year2}}))
  
  
  prop.test(x=pairwise_years$Confirmed,
            n=pairwise_years$Tested,
            alternative = c("two.sided"),
            conf.level = 0.95, correct = TRUE)}
yearproptest("genesee", 2012, 2016)
yearproptest("genesee", 2012, 2014)
yearproptest("genesee", 2012, 2013)

```


```{r county tibble}
countylist <- c("Genesee","Saginaw")
countypropestimates <- c(0.015,0.008)
tibble(countylist,countypropestimates)
```

## Results Between Years

-2013 had significantly **lower** levels than all other years.

-2012 had significantly **higher** levels than 2014.
```{r year tibble}
yearlist <- c(2012,2014)
yearpropestimates <- c(0.017, 0.012)
tibble(yearlist,yearpropestimates)
```

## Limitations
-Data only spanned from 2012-2016.

-Flint's conditions weren't even considered an emergency until 2016 and wasn't declared safe to drink again until 2019.

-Possible limited access to lead testing.

-CDC changed elevated standard to 3.5 µg/dL.

## Links to Sources
https://rstudio.github.io/DT/

https://jtr13.github.io/cc19/different-ways-of-plotting-u-s-map-in-r.html

https://ggplot2-book.org/maps.html

https://www.consumernotice.org/environmental/water-contamination/flint-michigan/

Data: https://www.cdc.gov/nceh/lead/data/state/midata.htm
